Outbreak response models of vaccine-preventable diseases in humans (1970-2019) - A systematic review (Supplementary Files)

### 2023-03-02 Epidemics_submission/outbreak_response_models_review_v2.tex

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first \cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% \cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to \cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Models and modelling practices for assessing the impact of outbreak response interventions to human vaccine-preventable diseases (1970-2019): A systematic review}
}
\newline
\\
James M. Azam\textsuperscript{1*},
Xiaoxi Pang\textsuperscript{2},
Elisha B. Are\textsuperscript{1, 3},
Juliet R.C. Pulliam\textsuperscript{1},
Matthew J. Ferrari\textsuperscript{4}
\\
\bigskip
\textbf{1} DSI-NRF Centre of Excellence in Epidemiological Modelling and Analysis (SACEMA), Stellenbosch University, Stellenbosch 7600, South Africa
\\
\textbf{2} Department of Mathematics, The University of Manchester, Manchester, UK
\\
\textbf{3} Department of Mathematics, Simon Fraser University, Burnaby, BC, Canada
\\
\textbf{4} Center for Infectious Disease Dynamics, Department of Biology, The Pennsylvania
State University, University Park, Pennsylvania, USA
\\
\bigskip

* jamesazam@sun.ac.za

\end{flushleft}
\section*{Abstract}
\subsection*{Background}
Mathematical modelling can aid outbreak response decision-making. However, this would require collaboration among model developers, decision-makers, and local experts to incorporate appropriate realism. We conducted a systematic review of modelling studies on human vaccine-preventable disease (VPD) outbreaks to identify patterns in modelling practices among collaborations. We complemented this with a mini review of eligible studies from the foot-and-mouth disease (FMD) literature.   
\subsection*{Methods}
Three databases were searched for studies published during 1970-2019 that applied models to assess the impact of an outbreak response. Per included study, we extracted data on author affiliation type (academic institution, governmental, and non-governmental organizations), whether at least one author was affiliated to the country studied, interventions, and model characteristics. Furthermore, the studies were grouped into two collaboration types: purely academic (papers with only academic affiliations), and mixed (all other combinations) to help investigate differences in modelling patterns between collaboration types in the human disease literature. Additionally, we compared modelling practices between the human VPD and FMD literature. 
\subsection*{Results}
Human VPDs formed 228 of 253 included studies. Purely academic collaborations dominated the human disease studies (56\%). Notably, mixed collaborations increased in the last seven years (2013 - 2019). Most studies had an author in the country studied (75.2\%) but this was more likely among the mixed collaborations. Contrasted to the human VPDs, mixed collaborations dominated the FMD literature (56\%). Furthermore, FMD studies more often had an author affiliated to the country studied (92\%) and used complex model design, including stochasticity, and model parametrization and validation. 
\subsection*{Conclusion}
The increase in mixed collaboration studies over the past seven years could suggest an increase in the uptake of modelling for outbreak response decision-making. We encourage more mixed collaborations between academic and non-academic institutions and the involvement of locally affiliated authors to help ensure that the studies suit local contexts.

\subsection*{Keywords}
Infectious disease dynamics; outbreak response modelling

\section*{Highlights}
\begin{enumerate}
	\item We explored the nature of collaborative modelling in the human-vaccine preventable disease outbreak response literature.
	\item Author collaborations were grouped into purely academic and mixed types.
	\item Purely academic collaborations dominated the human vaccine-preventable diseases literature.
	\item Mixed collaborations were more likely to involve local experts and use complex methods.
	\item Mixed collaborations have more potential to lead to decision-making and are recommended.
\end{enumerate}

%\subsection*{Registration}
%A protocol following the PRISMA guidelines for systematic review protocols and outlining the procedures for this systematic review was registered on PROSPERO with registration number CRD42020160803 and published through a peer-reviewed process. 
\linenumbers

\section*{Introduction}
Successful outbreak response to infectious diseases is often the result of a highly collaborative process~\cite{Sigfrid2020}. Decision-making during outbreak response usually requires collaborations between academic and field experts, including governmental and Non-Governmental Organizations (NGOs)~\cite{Kretzschmar2020,Kerkhove2012,Okiror2021,Whitty2014,Sigfrid2020}. These collaborations ensure that important perspectives from both research and operations/implementation are accounted for in the decision-making process.

Outbreak response decision-making often requires adaptations to the affected location ~\cite{Whitty2014,Heesterbeek2015,Sigfrid2020}. The interventions and decisions are usually driven by data/evidence, information, and experiences from the past or in real-time, either locally or from similar phenomena elsewhere ~\cite{Abramowitz2015,Kerkhove2012}. Furthermore, involving local experts in the design of interventions has been found to boost the success of outbreak response efforts ~\cite{Sigfrid2020}. Consequently, outbreak response teams should ideally involve at least one local expert to provide more context~\cite{Sigfrid2020}.

Mathematical modelling can support outbreak response decision-making~\cite{Lofgren2015}. Modelling is a proven tool for revealing insights about the extent of disease spread, and impact of interventions, while drawing on lessons learnt to provide recommendations for decision-making during outbreaks~\cite{Kretzschmar2020,Whitty2014,Lofgren2015}. Insights from mathematical modelling, though often useful, only form part of the larger context (socio-economic, political, and so forth) to be considered during an outbreak, making it difficult to determine the extent to which it contributes to outbreak response decision-making ~\cite{Kretzschmar2020,Kerkhove2012}. One way to help ensure that modelling contributes to decision-making could be through the conduct of interdisciplinary research between model developers and decision-makers.

The link between interdisciplinarity in scientific research --- that is, research conducted by authors with diverse scientific backgrounds --- and research impact, for example, number of citations has been well-researched~\cite{Abramo2017,Wagner2011}. However, few studies have investigated the impact of interdisciplinary collaborations on the conduct of scientific research. One such study investigated the impact of interdisciplinarity on the scientific validity of the methods used in a selection of papers that applied machine learning on topics in biology or medicine~\cite{Littmann2020}. The study found that the methods used in the reviewed studies differed according to the nature of the scientific backgrounds of the authors who conducted the research. Our preliminary searches of the literature did not reveal any of such review studies investigating differences in research practices among collaborations in the outbreak response modelling literature. We, therefore, conducted a systematic review of outbreak response modelling studies of human vaccine-preventable diseases with the goal to explore the time and geographic (countries) patterns of collaboration types, and to ascertain differences in practices between the collaboration types in terms of the choice of model characteristics, modelling methods, and other topics.

We sought to answer the following questions with respect to the collaboration landscape in the human disease outbreak response modelling literature: (1) How are the studies distributed in time and geographic locations, and how often is at least one of the collaborators geographically connected to the location being studied? (2) What types of interventions are generally modelled and is there a difference between the conclusions drawn about the impact of interventions, especially, vaccination in comparison with other non-vaccination interventions during outbreaks? (3) Do the model characteristics (structure and dynamics) and modelling practices (parametrization, validation, and so forth) differ between the collaboration types?

We acknowledge that there is a parallel body of relevant modelling work in the livestock literature. Specifically, the outbreak response effort that resulted from the 2001 foot-and-mouth disease (FMD) virus outbreak in the United Kingdom has been a foundational example of the application of mathematical modelling for decision-making. Subsequent work based on this outbreak has been instrumental in the development of FMD outbreak intervention strategies around the world. Hence, even though the primary scope of this review is the application of models to inform outbreak response for human infectious diseases, we include the relevant FMD references for comparison. In the end, we, therefore, study the differences in terms of collaborations in the human disease modelling literature, and compare our observations to that of the FMD modelling literature to further ascertain any differences in approaches and practices between the human and livestock literature.

%The following were the objectives of the systematic review:
%\begin{enumerate}
%	\item To ascertain differences in model characteristics and modelling methods between collaboration types. 
%	\item To determine the distribution of outbreak response modelling studies over time (years) overall and in terms of collaboration types, 
%	\item To determine how often at least one of the authors on a paper was geographically connected to the country/location studied and whether there were differences between collaboration types,
%	\item To investigate types of interventions modelled and differences between the conclusions drawn about the impact of interventions.
%\end{enumerate}

\section*{Materials and methods}
We followed the 2020 Preferred Reporting Items for Systematic Reviews and Meta-analyses statement (PRISMA 2020)  to conduct this systematic review \cite{Page2021}.

\subsection*{Eligibility criteria}
The following definitions were used in the determining study eligibility:

\begin{itemize}
	\item ``Outbreak'': a new and sudden rise in the number of cases of a disease in a population, which when left uncontrolled, could lead to large scale geographic spread, AND 
	\item ``Outbreak response'': an intervention directly triggered by the outbreak of an infectious disease, AND
	\item ``Mechanistic models'': mathematical models that use an equation or system of equations to capture the biological mechanisms driving the transmission dynamics of the infectious disease at the level of an individual or population \cite{Lessler2016a,Reiner2013}, AND
	\item ``Outbreak intervention assessment'': a mechanistic model-based evaluation of the impact of an outbreak response intervention.
\end{itemize}

We included studies that used a mechanistic model to investigate the impact of interventions triggered by the outbreak of any of the human vaccine-preventable diseases we considered (Table 1 in \nameref{S1_File}), Ebola, or foot-and-mouth disease. Studies were considered unique even if they had a duplicated author list or a slight variation in the author list, probably representing the same modelling group, so far as the content of the paper was different.

We excluded studies that satisfied at least one of the following criteria:

\begin{itemize}
	\item Study used a model that is not mechanistic,
	\item Study is not about an outbreak as defined,
	\item Disease studied is not a human vaccine-preventable disease (Table 1 in \nameref{S1_File}), Ebola or foot-and-mouth disease,
	\item Study's objective is not to evaluate the impact of an intervention mounted in response to an outbreak, or
	\item Study is not published in English.
\end{itemize}

\subsection*{Information sources}
On January 15, 2020, we searched Scopus, PubMed, and Web of Science for eligible records. We searched each database from its earliest date of coverage through January 15, 2020.

\subsection*{Search strategy}
We constructed and validated search strings specific to each database. To validate the search strings, we first ran them in the specific database's search engine and obtained a database of records. We then searched the database for well-known papers that fit the criteria for included studies. We did this for each disease on our list.

The search strings were constructed by all five reviewers (JMA, XP, EBA, MJF, and JRCP) and in consultation with the Faculty of Science Librarian of Stellenbosch University to reflect three main topics and their synonyms, that is, ``outbreak'', ``intervention'', and ``mechanistic model''.

The following are the search strings per database.

\textbf{Scopus (Title, abstract, keywords search)}:

( TITLE-ABS-KEY ( epidemic  OR  outbreak  OR  emergency  OR  reactive  OR  crisis ) )  AND  ( TITLE-ABS-KEY ( respon*  OR  manage*  OR  control  OR  interven*  OR  strateg* ) )  AND  ( TITLE-ABS-KEY ( stochastic  OR  transmission  OR  computational  OR  mathematical  OR  mechanistic  OR  statistical  OR  simulation  OR  "In silico"  OR  dynamic* ) )  AND  ( TITLE-ABS-KEY ( model* ) )  AND  ( ( TITLE-ABS-KEY ( cholera  OR  dengue  OR  diphtheria  OR  ebola  OR  "Foot-and-mouth"  OR  "foot and mouth"  OR  fmd  OR  "Hepatitis A"  OR  "Hepatitis B"  OR  "Hepatitis E"  OR  "Haemophilus influenzae type b"  OR  hib  OR  "Human papillomavirus"  OR  hpv  OR  influenza ) )  OR  ( TITLE-ABS-KEY ( "Japanese encephalitis"  OR  malaria  OR  measles  OR  "Meningococcal meningitis"  OR  mumps  OR  pertussis  OR  "Whooping cough"  OR  "Pneumococcal disease"  OR  poliomyelitis  OR  polio  OR  rabies  OR  rotavirus  OR  rubella ) )  OR  ( TITLE-ABS-KEY ( tetanus  OR  "Tick-borne encephalitis"  OR  tuberculosis  OR  typhoid  OR  varicella  OR  chickenpox  OR  "Yellow Fever"  OR  "vaccine-preventable" ) ) )

\textbf{PubMed (Title and abstract search)}:

### 2023-03-02 analyses/scripts/full_data_extraction_citation_table.Rmd
moved analysis scripts and data into a folder

---
title: "Analysis of the included studies"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
    keep_tex: yes
    fig_crop: no
---

```{r echo=FALSE, include=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('ggthemes')
library('mdthemes')
library('scales')
library('janitor')
library('forcats')
library('stringr')
library('countrycode')
library('bib2df')
library('conflicted')
library('Hmisc')
library('ggwordcloud')

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'jpeg')
knitr::opts_chunk$set(out.height = "\\textheight",  out.width = "\\textwidth")
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

@@

#Rearrange the columns
compact_data_with_citation_keys <- compact_data_with_citation_keys %>% 
  relocate(c(title, bibtexkey), .before = publication_year) %>% 
  relocate(publication_year, .before = bibtexkey) %>% 
  relocate(c(x0_reviewer, category, author:language), 
           .after = collab_type
           )

# Remove the FMD papers (NB: We only use the FMD papers for comparison)
long_data_no_fmd <- long_data %>% filter(disease != 'fmd')

compact_data_with_citation_keys_no_fmd <- compact_data_with_citation_keys %>% 
  filter(disease != 'fmd')

#Filter the FMD data
fmd_data <- compact_data_with_citation_keys %>% 
  filter(disease == 'fmd') 
```

# Human vaccine-preventable diseases

```{r eval=FALSE}
library(PRISMAstatement)

prisma_chart <- prisma(found = 12986,
       found_other = 0,
       no_dupes = 5012, 
       extra_dupes_box = TRUE,
       screened = 5012, 
       screen_exclusions = 4211, 
       full_text = 801,
       full_text_exclusions = 548, 
       qualitative = 253, 
       quantitative = 0,
       width = 800, height = 800,
       font_size = 12)

prisma_chart
```

<!-- ## Publications per year -->

<!-- ```{r} -->

<!-- publications_per_year <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(publication_year)  -->

<!-- publications_per_year_plot <- ggplot(data = publications_per_year) +  -->
<!--   geom_bar(aes(x = publication_year,  -->
<!--                y = n -->
<!--                ),  -->
<!--            stat = 'identity',  -->
<!--            color = 'black', -->
<!--            fill = 'tomato3' -->
<!--            ) + -->
<!--   scale_y_continuous(breaks = seq(0, 30, 2), -->
<!--                      labels = seq(0, 30, 2) -->
<!--                      ) + -->
<!--   labs(x = '**Year of publication**', y = '**Number of papers**') + -->
<!--   md_theme_minimal(base_size = 12) -->

<!-- plot(publications_per_year_plot) -->

<!-- ggsave(filename = 'publications_per_year.tiff',  -->
<!--        plot = publications_per_year_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_fig_here -->
<!--        ) -->

<!-- ggsave(filename = 'publications_per_year.tiff',  -->
<!--        plot = publications_per_year_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_thesis_fig_here -->
<!--        ) -->

<!-- ``` -->

<!-- ## Collaboration patterns in time -->

<!-- ### Number of studies per combinations of author affiliation types -->

<!-- ```{r echo=FALSE} -->
<!-- #' Count by collaboration types and make a bar plot -->
<!-- collab_types_count <- compact_data_with_citation_keys_no_fmd %>% -->
<!--     count(author_affiliation_type, sort = TRUE) %>%  -->
<!--     arrange(desc(n)) %>%  -->
<!--     adorn_totals() %>%  -->
<!--     adorn_percentages(denominator = 'col') %>%  -->
<!--     adorn_pct_formatting() %>%  -->
<!--     adorn_ns() -->

<!-- knitr::kable(collab_types_count,  -->
<!--              format = "latex", booktabs = TRUE,  -->
<!--              caption = 'Number of publications per author affiliation type combination') %>%  -->
<!--         kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))  -->
<!-- ``` -->

<!-- ### Proportions of the total publications per year by collaboration type -->

<!-- ```{r echo=FALSE, fig.width=8} -->
<!-- #' Count the number of each author affiliation type per year (including FMD) -->
<!-- collap_types_by_year <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(year_aggreg,  -->
<!--         collab_type,  -->
<!--         name = 'total_publications' -->
<!--         ) -->

<!-- #The plot -->
<!-- # Modify the x-axis labels to account for the lumping of the years from 2005 and below -->
<!-- x_axis_labels <- as.character(seq(2005, 2020, 2)) -->

<!-- x_axis_labels[1] <- '1970-2005' -->

<!-- collab_proportions_plot <- ggplot(data = collap_types_by_year) +  -->
<!--   geom_bar(aes(x = year_aggreg, -->
<!--                y = total_publications, -->
<!--                fill = collab_type -->
<!--                ), -->
<!--            color = 'black', -->
<!--            stat = 'identity', -->
<!--            position = position_fill(reverse = TRUE)#, -->
<!--            #show.legend = FALSE -->
<!--              ) + -->
<!--   scale_fill_manual(values = c('tan3', 'turquoise2')) + -->
<!--   scale_x_continuous(breaks = seq(2005, 2020, 2), -->
<!--                      labels = x_axis_labels, -->
<!--                      expand = c(0, 0) -->
<!--                      ) + -->
<!--   #coord_flip() + -->
<!--   labs(#title = 'Publications per collaboration type over time', -->
<!--        x = '**Year**', -->
<!--        y = '**Proportion of total publications**', -->
<!--        fill = '**Collaboration type**'  -->
<!--          ) + -->
<!--   theme(legend.position = 'none') + -->
<!--  # md_theme_minimal(base_size = 16) #For Powerpoint -->
<!--   md_theme_minimal(base_size = 12) -->

<!-- plot(collab_proportions_plot) -->

<!-- ggsave(filename = 'collab_proportions.tiff',  -->
<!--        plot = collab_proportions_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_fig_here -->
<!--        ) -->

<!-- ggsave(filename = 'collab_proportions.tiff',  -->
<!--        plot = collab_proportions_plot,  -->
<!--        device = 'tiff',  -->
<!--        path = save_thesis_fig_here -->
<!--        ) -->

<!-- ``` -->

<!-- ### Absolute number of publications per year by collaboration type -->

<!-- ```{r} -->
<!-- #' Count the number of each author affiliation type per year -->
<!-- collabs_per_year <- compact_data_with_citation_keys_no_fmd %>%  -->
<!--   count(year_aggreg,  -->
<!--         collab_type,  -->
<!--         name = 'total_publications' -->
<!--         )  -->

<!-- collabs_per_year_plot <- ggplot(data = collabs_per_year,  -->
<!--                                     aes(x = year_aggreg,  -->
<!--                                         y = total_publications -->
<!--                                         ),  -->
<!--                                     color = 'black') +  -->
<!--   geom_bar(aes(fill = collab_type),  -->
<!--            stat = 'identity',  -->
<!--            position = 'dodge' -->
<!--            ) +  -->
<!--   scale_fill_manual(values = c('tan3', 'turquoise2')) +  -->
<!--   scale_x_continuous(breaks = seq(2005, 2020, 2), -->
<!--                      labels = x_axis_labels, -->
<!--                      expand = c(0, 0) -->
<!--                      ) + -->
<!--   labs(x = '**Year**',  -->
<!--        y = '**Total publications**',  -->
<!--        fill = 'Collaboration type' -->
<!--        ) + -->
<!--   md_theme_minimal(base_size = 12) + -->
<!--   #md_theme_minimal(base_size = 16) + -->
<!--   theme(legend.position = 'bottom')  -->

<!-- plot(collabs_per_year_plot) -->

<!-- ggsave(filename = 'total_collabs_per_year_plot.tiff', -->
<!--        plot = collabs_per_year_plot,  -->
<!--        device = 'tiff', -->
<!--        path = save_fig_here -->
<!--        ) -->

<!-- ggsave(filename = 'total_collabs_per_year_plot.tiff', -->
<!--        plot = collabs_per_year_plot,  -->
<!--        device = 'tiff', -->
<!--        path = save_thesis_fig_here -->
<!--        ) -->

<!-- ``` -->

<!-- Even though non-academic collaborations increased over time, the share of total publications per year over time decreased. -->

<!-- ## Collaboration patterns in geographic space -->

## Collaboration types (aggregated)
```{r}
#' Count by collaboration types and make a bar plot
studies_per_collab_type <- compact_data_with_citation_keys_no_fmd %>%
    count(collab_type, sort = TRUE) %>% 
    arrange(desc(n)) %>% 
    adorn_totals() %>% 
    adorn_percentages(denominator = 'col') %>% 
    adorn_pct_formatting() %>% 
    adorn_ns()

knitr::kable(studies_per_collab_type, caption = 'Number of studies per collaboration type (Human VPDs)',
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Locations studied

### Real versus hypothetical

```{r}
location_types_tally <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>%
    mutate(location_type = ifelse(country_studied == 'none', 'none', 'real')) %>% 
    count(location_type) %>% 
  arrange(desc(n)) %>% 
   adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(location_types_tally, 
             caption = 'Number of studies per location type', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

### Larger than one country

```{r}
larger_geographies_tally <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>%
    filter(country_studied %in% c('west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
    count(country_studied) %>% 
  arrange(desc(n)) %>% 
   adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

knitr::kable(larger_geographies_tally, caption = 'Number of studies per location larger than one country', format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))

```

### Countries studied

```{r}
countries_studied_by_count <- compact_data_with_citation_keys_no_fmd %>% 
    separate_rows(country_studied, sep = ' ') %>%
    distinct(title, country_studied, .keep_all = TRUE) %>% 
    count(collab_type, country_studied, name = 'num_of_studies', sort = T)

#'Modify the data frame to lump together all countries with less than some number of studies as 'other' and recalculate the percentage studied
countries_studied_by_count_aggregated <- countries_studied_by_count %>% 
  group_by(country_studied) %>%
  summarise(num_of_studies = sum(num_of_studies),
            .groups = 'drop'
            ) %>% 
  filter(country_studied %nin% c('none', 'west_africa', 'global', 'northern_hemisphere', 'southeast_asia', 'who_southeast_asia_region')) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  ungroup()

#Print the results
knitr::kable(countries_studied_by_count_aggregated, 
             caption = 'Number of studies per country', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'), font_size = 7)
```

### Continents studied

```{r}
continents_studied_by_count <- countries_studied_by_count %>% 
  filter(country_studied %nin% c('none', 'northern_hemisphere', 'global')) %>% 
  mutate(continent_studied = countrycode(country_studied, 
                                          destination = 'continent',
                                          origin = 'iso2c',
                                          nomatch = NULL #if there is no match, keep the original
                                          )
         ) %>% 
  mutate(continent_studied = str_to_lower(continent_studied)
         ) %>% 
  mutate(continent_studied = case_when(str_detect(continent_studied, 'asia') == T ~ 'asia', 
                                         str_detect(continent_studied, 'africa') == T ~ 'africa', 
                                         TRUE ~ continent_studied
                                       )
           ) %>% 
  group_by(continent_studied) %>%
  summarise(num_of_studies = sum(num_of_studies), 
            .groups = 'drop'
            ) %>% 
  arrange(desc(num_of_studies)) %>% 
  adorn_totals(where = c('row')) %>% 
  adorn_percentages(denominator = 'col') %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

#Print the results
knitr::kable(continents_studied_by_count, 
             caption = 'Number of studies per continent', 
             format = "latex", booktabs = TRUE) %>% 
        kableExtra::kable_styling(latex_options = c('striped', 'HOLD_position'))
```

## Geographic connection of authors to the studied locations

>>>>

---
output:
  pdf_document:
    keep_tex: yes
---

```{r echo=FALSE, include=FALSE}
#' load packages
library('tidyverse')
library('readr')
library('xlsx')
library('scales')
library('stringr')
library('conflicted')
library('Hmisc')
library('kableExtra')

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'jpeg')
knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
knitr::opts_chunk$set(message = FALSE)

@@

#Rearrange the columns
compact_data_with_citation_keys <- compact_data_with_citation_keys %>% 
  relocate(c(title, bibtexkey), .before = publication_year) %>% 
  relocate(publication_year, .before = bibtexkey) %>% 
  relocate(c(x0_reviewer, category, author:language), 
           .after = collab_type
           )
```

```{r results='asis'}
final_citation_table <- compact_data_with_citation_keys %>% 
    relocate(c('doi', 'year', 'bibtexkey'), .after = title) %>% 
    select(-c(year_aggreg, x0_reviewer, category, 
              number, pages, journal, publisher, 
              volume, language, month, publication_type, 
              modelling_timing, intervention_type, collab_type,
              author, paper_doi, year, 
              doi, bibtexkey
              )
           ) %>% 
    arrange(desc(publication_year))
```

```{r}
#columns 2-5 with paper titles
knitr::kable(final_citation_table %>% 
                 dplyr::select(1:5), 
             format = "latex", booktabs = TRUE, 
             longtable = TRUE#,
             #caption = 'Extracted data per paper: publication_year, author_affiliation_type, country_studied, author_in_country_studied.'
             ) %>% 
    kable_styling(latex_options = c('striped', 'scale_down', 'hold_position', 'repeat_header')) %>% 
    kableExtra::column_spec(column = c(1, 3), width = "5cm") %>% 
    kableExtra::landscape()
```

```{r}
#columns 6-10 with paper titles
knitr::kable(final_citation_table %>% 
                 dplyr::select(1, 6:10), 
             format = "latex", booktabs = TRUE, 
             longtable = TRUE#,
             #caption = 'Extracted data per paper: disease, objectives, individual_representation, space_representation, model_structure.'
             ) %>% 
    kable_styling(latex_options = c('striped', 'scale_down', 'hold_position', 'repeat_header')) %>% 
    kableExtra::column_spec(column = c(1), width = "5cm") %>% 
    kableExtra::landscape()
```

```{r}
#columns 11-15 with paper titles
knitr::kable(final_citation_table %>% 
                 dplyr::select(1, 11:15), 
             format = "latex", booktabs = TRUE, 
             longtable = TRUE#,
           #  caption = 'Extracted data per paper: parametrization, validation, outbreak_type, intervention_modelled, vax_modelled_with_non_vax.'
             ) %>% 
    kable_styling(latex_options = c('striped', 'scale_down', 'hold_position', 'repeat_header')) %>% 
    kableExtra::column_spec(column = c(1, 2, 5), width = "4cm") %>% 
    kableExtra::landscape()
```

```{r}
#columns 16-22 with paper titles
knitr::kable(final_citation_table %>% 
                 dplyr::select(1, 16:22), 
             format = "latex", booktabs = TRUE, 
             longtable = TRUE#,
           #  caption = 'Extracted data per paper: is_vax_effective, outcome_measured, sensitivity_analysis, data_used, data_available, simulation_code_available, vax_modelled.'
             ) %>% 
    kable_styling(latex_options = c('striped', 'scale_down', 'hold_position', 'repeat_header')) %>% 
    kableExtra::column_spec(column = c(1, 3), width = "5cm") %>% 
    kableExtra::landscape()
```
